TacticToe: Learning the Game of Tactics

### 2018-01-17 tactictoe_jv.tex
AITP integration: wip

A large difference between diagonalized search and our MCTS algorithm is in 
what is considered to be the width of the tree. In diagonalized search, 
non-productive tactics contribute to the extension of the width where as in 
MCTSn the width includes only the total number of children produced. The 
influence of non-productive on the proof exploration is recovered when we 
include an 
evaluation function that takes into account the productivity of tactics such as 
$E_{productive}$ or $E_r$.

@@

\section{ATP Integration}\todo{JU + CK: rework and review}

General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
\holyhammer for \holfour~\cite{tgck-cpp15}. The principle is that we will 
create a \holyhammer tactic that can be used during the proof search. Since 
external calls to \holyhammer are computationally expensive (5 seconds at a 
minimum), we  
experiment with a small hammer tactic comprised of: a faster premise 
selection algorithm from previously proven theorems $\mathfrak{T}$ and
a short call (0.1 seconds + 16 premises optimized by the previous paper) to the 
internal prover \metis~\cite{metis} or a
longer call to external prover \eprover.

We construct this type of tactic manually but in the case of \metis it could be 
rediscovered automatically as it is essentially a call to \metis where its 
theorem list argument has been abstracted.

Since we do not always have the data to learn on which goal this general 
tactics are succesful, we assume as a first approximation that they are the 
strongest all around. And we try them first on each node during the proof 
search with the advantage that if it succeeds the node is closed rapidly in the 
case of \metis.

To integrate a general tactic that requires a higher time out such as \eprover, 
we make use of asynchronous calls to prevent the ATP 
from slowing down the search.  The number of parallel threads can be capped.
And if tactictoe or an asynchronous thread proves a goal, all descendants of 
these goal are deactivated, terminating associated threads. This frees space 
for more pending threads, associated to newly created nodes, to start.

>>>>

%A large difference between diagonalized search and our MCTS algorithm is in 
%what is considered to be the width of the tree. In diagonalized search, 
%non-productive tactics contribute to the extension of the width where as in 
%MCTSn the width includes only the total number of children produced. The 
%influence of non-productive on the proof exploration is recovered when we 
%include an 
%evaluation function that takes into account the productivity of tactics such 
%as 
%$E_{productive}$ or $E_r$.

@@

\section{ATP Integration}\todo{TG + JU + CK: rework and review}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
stronger general purpose provers such as the one in \holyhammer for 
\holfour~\cite{tgck-cpp15}.

During our experiments,
We chose to distinguish two these general purpose tactic \metis and \eprover.
\metis is integrated in \holfour, so can be call with frequently with low 
timeout.
And the strongest external provers according to past experiments \cite{hh4h4}.
Where as...
external calls to \eprover are computationally expensive (5 seconds at a 
minimum +  128 premises), we  
experiment with a small hammer tactic comprised of: a faster premise 
selection algorithm and
a short call (0.1 seconds + 16 premises optimized by the previous paper) to the 
internal prover \metis~\cite{metis} or a
longer call to external prover \eprover.

\paragraph{Prediction of general purpose tactics}
Since we do not always have the data to learn on which goal this general 
tactics are successful, we assume as a first approximation that they are the 
strongest all around.

And we try them first on each node during the proof search with the advantage 
that if it succeeds the node is closed rapidly.

This general tactics requires a list of relevant lemmas.

This forced behavior could often be rediscovered from the training data for the 
internal provers \metis by using theorem abstraction. essentially a call to 
\metis where its 
theorem list argument has been abstracted.

\paragraph{Application of general purpose tactics}
The slightly higher timeout of 0.1 seconds given to \metis does not slowdown 
the proof too much but...
To integrate a general tactic that requires a higher time out such as \eprover, 
we make use of asynchronous calls via PolyML.fork to prevent the ATP 
from slowing down the search.

%The number of parallel threads can be capped.
%And if tactictoe or an asynchronous thread proves a goal, all descendants of 
%these goal are deactivated, terminating associated threads. This frees space 
%for more pending threads, associated to newly created nodes, to start.
